---

This code aims to give a comparision between weekdays and weekend average steps were taken on every 5 minitues intervals.
-----------------------------------------------------------------------------------------

Code for reading in the dataset and/or processing the data
This  statement is used to load the data by read.csv().

```{r}
daily_exercise <-
  read.csv("C:/Users/Arash/Documents/Coursera/Reproducible Research/activity.csv", header = TRUE, sep = ",")
```
===============================================================================

Convert the date into date data type and interva field to Factor class

Using Lubridate provides tools that make it easier to parse and manipulate dates.

```{r}
daily_exercise$date <- as.Date(daily_exercise$date, format = "%Y-%m-%d")
```
================================================================================

What is mean total number of steps taken per day?

```{r}

library(dplyr)

 Perday <- daily_exercise %>% group_by(date) %>%
 
  summarize(total_steps = sum(steps, na.rm = T),
           mean_steps = mean(steps, na.rm = T))
 ```
 
 ================================================================================
 
 make a histogram of the total number of steps taken each day
               
  ```{r}
  library(ggplot2)
  g<- ggplot(Perday, aes(x=total_steps))
 g + geom_histogram(binwidth = 2500) + theme(axis.text = element_text(size = 12),  
     axis.title = element_text(size = 14)) + labs(y = "Frequency") + labs(x = "Total steps/day")    
     ```
     
===================================================================================

Mean and median number of steps taken each day

Calculate and report the mean and median of the total number of steps taken per day
 
```{r}
summary(Perday$total_steps)
 
 summary(Perday$mean_steps)
 ```
==================================================================================== 
     
 What is the average daily activity pattern?     
 
 ```{r}
 
  Interval_5_minutes<- daily_exercise %>% group_by(interval) %>%
     summarize(mean_steps = mean(steps, na.rm = T))
 

  g <- ggplot(Interval_5_minutes, aes(x = interval, y = mean_steps))
 g + geom_line() + theme(axis.text = element_text(size = 12),  
                                             axis.title = element_text(size = 14))  +
   labs(y = "mean number of steps") + labs(x = "Interval")
   ```
=====================================================================================


The following lines will calculate the percentage of missing data as 
well as the number of rows that contain an NA.

```{r}

mean(is.na(daily_exercise$steps))
sum(is.na(daily_exercise$steps))

```
=====================================================================================

locating NAs and replacing these missing values

There are several alternatives we can use to fill the NAs
Using the average steps per interval will be the approach to fill in the NAs

since There are no missing values in this variable so I will use it to fill in for NAs.

```{r}

 sum(is.na(Interval_5_minutes$mean_steps))
```

Creating fresh copy of our original data

```{r}

NewDb <- daily_exercise
```
before filling the NAs in our data, check to see which rows has NAs in steps column

Using the following loop to find and replace NAs

```{r}

for (i in 1:nrow(NewDb)) {
      if (is.na(NewDb$steps[i])) {
          k <- NewDb$interval[i]
          v <- subset(Interval_5_minutes, interval == k)
          NewDb$steps[i] <- v$mean_steps
      }
 }
 
 ```
=============================================================================

 In this section I caclualte the total daily steps
 
 ```{r}
 
 newPerday <- NewDb %>% group_by(date) %>%
   
   summarize(total_steps = sum(steps, na.rm = T),
             mean_steps = mean(steps, na.rm = T))
 
 # New Histogram with the new Data without NAs
 
 g<- ggplot(newPerday, aes(x=total_steps))
 g + geom_histogram(binwidth = 2500) + theme(axis.text = element_text(size = 12),  
                                             axis.title = element_text(size = 14)) + labs(y = "Frequency") + labs(x = "Total steps/day")                    
 ```
================================================================================

summary and standard deviation of the new data 

```{r}

 summary(newPerday$total_steps)
 sd(newPerday$total_steps, na.rm = T)
 ```
 ==============================================================================
 
 comparing the summary and standard deviaiton of new data to original data
 
 ```{r}
 
 summary(Perday$total_steps)
 sd(Perday$total_steps, na.rm = T)
 
 ```
 
 smaller standard deviaiton of new data without NAs shows that repalcing NAs with mean
 values for the time intervals decrased the spread
  ===============================================================================
 
 Panel plot comparing the average number of steps taken per 5-minute interval 
 across weekdays and weekends
 
 
 creating a new column containing weekend and weekdays values in Newdb
 
 ```{r}
 
 NewDb$day <- ifelse(weekdays(NewDb$date) %in% c("Saturday","Sunday"), "weekend","weekday")
 ```
 
 creating two subsets , one for weekdays data and one for weekend data
 ```{r}
 
  weekday <- filter(NewDb, day == "weekday")
  weekend <- filter(NewDb, day == "weekend")
  
  ```
  Again I group these two data by interval and I calculate the mean number
  of steps for each interval
  
  ```{r}
  
   weekday <- weekday %>%
             group_by(interval) %>%
              summarize(mean_steps = mean(steps))
              weekday$day <- "weekday"
  
  weekend <- weekend %>%
              group_by(interval) %>%
              summarize(mean_steps = mean(steps))
              weekend$day <- "weekend"
              
```

now I merge these two data into one data set

```{r}

NewInterval_5_minutes <- rbind(weekday, weekend)
    NewInterval_5_minutes$day <- as.factor(NewInterval_5_minutes$day)
    NewInterval_5_minutes$day <- relevel(NewInterval_5_minutes$day, "weekend")
    
    
    
    for (j in 1:nrow(NewInterval_5_minutes)) {
      if (is.na(NewInterval_5_minutes$mean_steps[j])) {
        kk <- NewInterval_5_minutes$interval[j]
        vv <- subset(Interval_5_minutes, interval == kk)
        NewInterval_5_minutes$mean_steps[j] <- vv$mean_steps
      }
  
    }
    
    ```

I plot weekdays and weekend  side by side by using day column as a factor to 
differentiate between weekdays and weekend    

```{r}

gg<- ggplot(NewInterval_5_minutes, aes(interval, mean_steps))
     gg + geom_line() +facet_grid(day~.) +theme(axis.text = element_text(12), axis.title = element_text(14)) +
      labs(y = "Number of Steps") +labs(x = "Interval")
    
    ```
    
    activity during weekdays and weekend is very different.
    during weekdays activity remains below 100 steps and peaks during morning
    in wekend there is no show of high level activity